---
title: "Supplementary Materials: Controls on litter decomposition at ambient and elevated CO2 concentrations"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

## load libraries
library(tidyverse)
library(ggeffects)
library(patchwork)
library(lme4)
library(car)
library(emmeans)
library(vegan)
library(rioja)
library(ggforce)
library(broom)
library(pixiedust)


## load and prep data

# soil parameters
soil <- read_csv('derivedData/soilvars.csv') %>% 
  mutate(Ring = as.factor(Ring)) %>% 
  rename(Time = Months)

# harvests
harvs <- bind_rows(data.frame(experiment = 'expt1',
                              Time = c(3, 6, 9, 12, 18, 24),
                              Date = as.Date(c('2013-08-23', '2013-12-04', '2014-03-13',
                                               '2014-06-16', '2014-11-12', '2015-03-19')),
                              Deployed = as.Date('2013-05-17')),
                   data.frame(experiment = 'expt2',
                              Time = 1:4,
                              Date = as.Date(c('2017-12-18', '2018-04-26', '2018-08-23', '2018-12-13')),
                              Deployed = as.Date('2017-08-25'))) %>%
  mutate(Months = as.numeric(round((Date - Deployed)/(365.25/12), 1))) %>%
  select(-Deployed)

# otu annotations
tax <- read_tsv('derivedData/otu_tax.tsv') %>% 
  filter(!is.na(phylum))

# expt 1, litter
litter_e1 <- read_csv('derivedData/litter_expt1.csv') %>% 
  mutate(across(c(Ring, Plot, Mesh), as.factor),
         Time.ord = ordered(Time)) %>%
  left_join(soil %>% 
              group_by(experiment, Time, Ring) %>% 
              summarise(VWC = mean(VWC, na.rm=TRUE), 
                        TDRTemp = mean(TDRTemp, na.rm=TRUE)))

# expt 1, chemistry and otus
otus_e1 <- read_csv('derivedData/otus_chem_expt1.csv') 

# expt 2, litter
litter_e2 <- read_csv('derivedData/litter_expt2.csv') %>% 
  mutate(across(c(Ring, Plot, Trt, prevTrt, psyllid), as.factor), 
         Time.ord = ordered(Time)) %>% 
  left_join(soil %>% 
              group_by(experiment, Time, Ring) %>% 
              summarise(VWC = mean(VWC, na.rm=TRUE), 
                        TDRTemp = mean(TDRTemp, na.rm=TRUE)))

# expt 2, chemistry and otus
otus_e2 <- read_csv('derivedData/otus_chem_expt2.csv')

```

\newpage

Table S1. Initial nitrogen and phosphorus concentrations of litter used in each decomposition study and, for the second study, each condition for leaf development prior to litter collection.

```{r}

# initial, pre-incubation
bind_rows(
  otus_e1 %>% 
    filter(Mesh == 'initial') %>% 
    summarise(study='study 1', prevTrt='aCO2', psyllid='noPsyl', n(), 
              N_mean = mean(N_percent), N_sd = sd(N_percent), 
              P_mean = mean(P_ppm)/10000, P_sd = sd(P_ppm)/10000),
  otus_e2 %>% 
    filter(Stage == 'initial') %>% 
    group_by(prevTrt, psyllid) %>% 
    summarise(study='study 2', n(), 
              N_mean = mean(N_percent), N_sd = sd(N_percent), 
              P_mean = mean(P_ppm)/10000, P_sd = sd(P_ppm)/10000)) %>% 
  mutate(across(starts_with('N_'), ~round(., 2)), 
         across(starts_with('P_'), ~round(., 3))) -> tmp
# tmp
# write_csv(tmp, 'output/SuppTable_initialChem')

dust(tmp) %>%
  sprinkle_colnames('Study', 'Previous CO2 condition', 'Psyllid exposure', 'Samples', 'Mean %N', 'SD %N', 'Mean %P', 'SD %P') %>%
  sprinkle(cols = 1, replace = c('1', '2', '', '', '')) %>%
  sprinkle(cols = 2, replace = c('ambient', 'ambient', '', 'elevated', '')) %>%
  sprinkle(cols = 3, replace = c('-', 'unimpacted', 'Psyllid-impacted', 'unimpacted', 'Psyllid-impacted')) %>%
  sprinkle(cols = 5:6, round = 1) %>% 
  sprinkle(cols = 7:8, round = 2) %>% 
  sprinkle_print_method("markdown")

```

```{r, eval=TRUE}

## this code chunk is not evaluated, no output is produced
## but this code estimates litter N and P responses and conducts statistical hypothesis tests 

# post-incubation but pre-effect, study 1
tmp <- otus_e1 %>% filter(Months == 6.6)
Anova(lmer(N_percent ~ Mesh * Trt + (1|Ring/Plot), data=tmp), test='F')
Anova(lmer(P_ppm ~ Mesh * Trt + (1|Ring/Plot), data=tmp), test='F')
tmp %>% 
  # group_by(Trt) %>% 
  summarise(N_mean = mean(N_percent), N_sd = sd(N_percent), 
            P_mean = mean(P_ppm)/10000, P_sd = sd(P_ppm)/10000)

# post-incubation but pre-effect, study 2
tmp <- otus_e2 %>% filter(Months == 8)
Anova(lmer(N_percent ~ Trt * prevTrt * psyllid + (1|Ring/Plot), data=tmp), test='F')
Anova(lmer(P_ppm ~ Trt * prevTrt * psyllid + (1|Ring/Plot), data=tmp), test='F')
tmp %>% 
  # group_by(Trt) %>% 
  summarise(N_mean = mean(N_percent, na.rm=TRUE), N_sd = sd(N_percent, na.rm=TRUE), 
            P_mean = mean(P_ppm, na.rm=TRUE)/10000, P_sd = sd(P_ppm, na.rm=TRUE)/10000)

```


\newpage

Table S2. PerMANOVA analysis of fungal assemblages in the first study (Trt = CO$_2$ treatment during decomposition, Mesh = hole size in litter-containing mesh). OTU tables were analysed using Bray-Curtis dissimilarities. 

```{r}

## analyse mesh/time and treatment effects (only samples prior to treatment effects on decomposition)
# NOTE THIS IS BASED ON PERMUTATION AND THERE WILL BE SOME VARIATION IN THE P-VALUES WITH EACH RUN
# THE RESULTS IN THE MANUSCRIPT ARE BASED ON THE FIRST TIME RUN

# prepare data
temp <- otus_e1 %>% 
  filter(Stage == 'preEffect') %>% 
  mutate(across(c(Trt, Mesh), as.factor))
mat <- decostand(as.matrix(temp %>% select(starts_with('ITSall_'))), 'hellinger')

# analysis
tmp <- adonis2(mat ~ Trt * Mesh, data=temp)
tmp

```

\newpage

Table S3. Indicator OTUs associated with ambient or elevated CO$_2$ conditions during litter decomposition in the first study.

```{r}

# study 1 - CO2 treatment
# NOTE THIS IS BASED ON PERMUTATION AND THERE WILL BE SOME VARIATION IN THE TAXA SELECTED WITH EACH RUN
# THE RESULTS IN THE MANUSCRIPT ARE BASED ON THE FIRST TIME RUN

indic <- indicspecies::multipatt(mat, temp$Trt, control=how(nperm=999))
inner_join(indic$A %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(ambient > 0.8 | elevated > 0.8) %>%
             select(OTU:elevated) %>%
             rename(A_ambient='ambient', A_elevated='elevated'),
           indic$B %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(ambient > 0.35 | elevated > 0.35) %>%
             select(OTU:elevated) %>%
             rename(B_ambient='ambient', B_elevated='elevated')) %>%
  inner_join(indic$sign %>%
               rownames_to_column('OTU') %>%
               filter(p.value <= 0.01) %>%
               mutate(group = case_when(s.ambient == 1 ~ 'aCO2',
                                        s.elevated == 1 ~ 'eCO2')) %>%
               select(OTU, group, stat, p.value) %>%
               arrange(group, desc(stat)) %>%
               left_join(tax) %>%
               select(OTU, group, stat, p.value, phylum, order, genus, species, guild, confidenceRanking)) %>% 
  arrange(group, desc(stat)) %>% 
  mutate(compA = case_when(group == 'aCO2' ~ A_ambient, 
                           group == 'eCO2' ~ A_elevated), 
         compB = case_when(group == 'aCO2' ~ B_ambient, 
                           group == 'eCO2' ~ B_elevated)) %>% 
  select(-starts_with('A_'), -starts_with('B_')) %>% 
  relocate(OTU, group, compA, compB, stat, p.value) -> tmp
dust(tmp) %>%
  sprinkle_colnames('OTU ID', 'Litter condition', 'A', 'B', 'Score', 'P-value', 'Phylum', 'Order', 'Genus', 'Species', 'Guild', 'Confidence ranking') %>%
  sprinkle(cols = 3:5, round = 2) %>%
  sprinkle(cols = 6, round = 3) %>%
  sprinkle_print_method("markdown")

```

\newpage

Table S4. PerMANOVA analysis of fungal assemblages in the second study (Trt = CO$_2$ treatment during litter decomposition, prevTrt = CO$_2$ treatment during leaf development, psyllid = whether litter was impacted by a psyllid outbreak during leaf development). OTU tables were analysed using Bray-Curtis dissimilarities. 

```{r}

## analyse mesh/time and treatment effects (only samples prior to treatment effects on decomposition)
# NOTE THIS IS BASED ON PERMUTATION AND THERE WILL BE SOME VARIATION IN THE P-VALUES WITH EACH RUN
# THE RESULTS IN THE MANUSCRIPT ARE BASED ON THE FIRST TIME RUN

# prepare data
temp <- otus_e2 %>% 
  filter(Stage == 'preEffect') %>% 
  mutate(across(c(Trt, prevTrt, psyllid), as.factor))
mat <- decostand(as.matrix(temp %>% select(starts_with('ITSall_'))), 'hellinger')

# analysis
tmp <- adonis2(mat ~ Trt * prevTrt * psyllid, data=temp)
tmp

```

\newpage

Table S5. Indicator OTUs associated with ambient or elevated CO$_2$ conditions during litter decomposition in the second study.

```{r}

indic <- indicspecies::multipatt(mat, temp$Trt, control=how(nperm=999))
inner_join(indic$A %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(aCO2 > 0.8 | eCO2 > 0.8) %>%
             select(OTU:eCO2) %>%
             rename(A_ambient='aCO2', A_elevated='eCO2'),
           indic$B %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(aCO2 > 0.35 | eCO2 > 0.35) %>%
             select(OTU:eCO2) %>%
             rename(B_ambient='aCO2', B_elevated='eCO2')) %>%
  inner_join(indic$sign %>%
               rownames_to_column('OTU') %>%
               filter(p.value <= 0.01) %>%
               mutate(group = case_when(s.aCO2 == 1 ~ 'aCO2',
                                        s.eCO2 == 1 ~ 'eCO2')) %>%
               select(OTU, group, stat, p.value) %>%
               arrange(group, desc(stat)) %>%
               left_join(tax) %>%
               select(OTU, group, stat, p.value, phylum, order, genus, species, guild, confidenceRanking)) %>% 
  arrange(group, desc(stat)) %>% 
  mutate(compA = case_when(group == 'aCO2' ~ A_ambient, 
                           group == 'eCO2' ~ A_elevated), 
         compB = case_when(group == 'aCO2' ~ B_ambient, 
                           group == 'eCO2' ~ B_elevated)) %>% 
  select(-starts_with('A_'), -starts_with('B_')) %>% 
  relocate(OTU, group, compA, compB, stat, p.value) -> tmp
dust(tmp) %>%
  sprinkle_colnames('OTU ID', 'Litter condition', 'A', 'B', 'Score', 'P-value', 'Phylum', 'Order', 'Genus', 'Species', 'Guild', 'Confidence ranking') %>%
  sprinkle(cols = 3:5, round = 2) %>%
  sprinkle(cols = 6, round = 3) %>%
  sprinkle_print_method("markdown")

```

\newpage

Table S6. Indicator OTUs associated with ambient or elevated CO$_2$ conditions during leaf development in the second study.

```{r}

# NOTE THIS IS BASED ON PERMUTATION AND THERE WILL BE SOME VARIATION IN THE TAXA SELECTED WITH EACH RUN
# THE RESULTS IN THE MANUSCRIPT ARE BASED ON THE FIRST TIME RUN

indic <- indicspecies::multipatt(mat, temp$prevTrt, control=how(nperm=999)) 
inner_join(indic$A %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(aCO2 > 0.8 | eCO2 > 0.8) %>%
             select(OTU:eCO2) %>%
             rename(A_ambient='aCO2', A_elevated='eCO2'),
           indic$B %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             # filter(aCO2 > 0.35 | eCO2 > 0.35) %>%
             select(OTU:eCO2) %>%
             rename(B_ambient='aCO2', B_elevated='eCO2')) %>%
  inner_join(indic$sign %>%
               rownames_to_column('OTU') %>%
               filter(p.value <= 0.01) %>%
               mutate(group = case_when(s.aCO2 == 1 ~ 'aCO2',
                                        s.eCO2 == 1 ~ 'eCO2')) %>%
               select(OTU, group, stat, p.value) %>%
               arrange(group, desc(stat)) %>%
               left_join(tax) %>%
               select(OTU, group, stat, p.value, phylum, order, genus, species, guild, confidenceRanking)) %>% 
  arrange(group, desc(stat)) %>% 
  mutate(compA = case_when(group == 'aCO2' ~ A_ambient, 
                           group == 'eCO2' ~ A_elevated), 
         compB = case_when(group == 'aCO2' ~ B_ambient, 
                           group == 'eCO2' ~ B_elevated)) %>% 
  select(-starts_with('A_'), -starts_with('B_')) %>% 
  relocate(OTU, group, compA, compB, stat, p.value) -> tmp
dust(tmp) %>%
  sprinkle_colnames('OTU ID', 'Leaf condition', 'A', 'B', 'Score', 'P-value', 'Phylum', 'Order', 'Genus', 'Species', 'Guild', 'Confidence ranking') %>%
  sprinkle(cols = 3:5, round = 2) %>%
  sprinkle(cols = 6, round = 3) %>%
  sprinkle_print_method("markdown")

```

\newpage

Table S7. Indicator OTUs associated with the timing of litter collection prior to initiation of the second study.

```{r}

# NOTE THIS IS BASED ON PERMUTATION AND THERE WILL BE SOME VARIATION IN THE TAXA SELECTED WITH EACH RUN
# THE RESULTS IN THE MANUSCRIPT ARE BASED ON THE FIRST TIME RUN

indic <- indicspecies::multipatt(mat, temp$psyllid)
inner_join(indic$A %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             select(OTU:Psyl) %>%
             rename(A_noPsyl='noPsyl', A_Psyl='Psyl'),
           indic$B %>%
             as.data.frame() %>%
             rownames_to_column('OTU') %>%
             select(OTU:Psyl) %>%
             rename(B_noPsyl='noPsyl', B_Psyl='Psyl')) %>%
  inner_join(indic$sign %>%
               rownames_to_column('OTU') %>%
               filter(p.value <= 0.01) %>% 
               mutate(group = case_when(s.noPsyl == 1 ~ 'unimpacted', 
                                        s.Psyl == 1 ~ 'psyllid-impacted')) %>% 
               select(OTU, group, stat, p.value) %>%
               arrange(group, desc(stat)) %>%
               left_join(tax) %>%
               select(OTU, group, stat, p.value, phylum, order, genus, species, guild, confidenceRanking)) %>% 
  arrange(group, desc(stat)) %>% 
  mutate(compA = case_when(group == 'unimpacted' ~ A_noPsyl, 
                           group == 'psyllid-impacted' ~ A_Psyl), 
         compB = case_when(group == 'unimpacted' ~ B_noPsyl, 
                           group == 'psyllid-impacted' ~ B_Psyl)) %>% 
  select(-starts_with('A_'), -starts_with('B_')) %>% 
  relocate(OTU, group, compA, compB, stat, p.value) -> tmp
dust(tmp) %>%
  sprinkle_colnames('OTU ID', 'Leaf condition', 'A', 'B', 'Score', 'P-value', 'Phylum', 'Order', 'Genus', 'Species', 'Guild', 'Confidence ranking') %>%
  sprinkle(cols = 3:5, round = 2) %>%
  sprinkle(cols = 6, round = 3) %>%
  sprinkle_print_method("markdown")

```

\newpage

```{r, fig.cap="Figure S1. Soil temperature and moisture conditions in the top 10cm of soil during the two studies. Lines represent average values across all sensors (48 for soil moisture, 16 for soil temperature). Dashed vertical lines indicate sampling dates.", fig.dim=c(7,6), dpi=300}

soil %>% 
  group_by(experiment, Date) %>% 
  summarise(VWC=mean(VWC, na.rm=TRUE), 
            TDRTemp=mean(TDRTemp, na.rm=TRUE)) %>% 
  pivot_longer(cols=c(VWC, TDRTemp), names_to='var', values_to='value') %>% 
  mutate(experiment=case_when(experiment == 'expt1' ~ 'First study', 
                              experiment == 'expt2' ~ 'Second study')) -> tmp

ggplot(tmp %>% filter(var=='TDRTemp'), aes(x=Date, y=value)) + 
  geom_line() + 
  labs(x='', y='Soil temperature (deg. C)') + 
  facet_grid(cols=vars(experiment), scales='free') + 
  geom_vline(xintercept=harvs$Date, linetype='dotted') -> p1
ggplot(tmp %>% filter(var=='VWC'), aes(x=Date, y=value)) + 
  geom_line() + 
  labs(x='', y='Soil volumetric water content (%)') + 
  facet_grid(cols=vars(experiment), scales='free') + 
  geom_vline(xintercept=harvs$Date, linetype='dotted') -> p2
p1 / p2

```

\newpage

```{r, fig.cap="Figure S2. Effect sizes associated with elevated CO2 on soil volumetric water content (VWC) were within the margin of error for almost the entire duration of both studies, with only one short period in the first study where soil moisture was lower in the elevated CO2 treatment relative to ambient conditions. Therefore, it is unlikely that differences in soil microclimate were responsible for treatment effects on decomposition rates. Solid lines represent mean effect sizes (negative values indicate drier soil moisture conditions under elevated CO2) and ribbons represent 95% confidence intervals. Dashed vertical lines indicate sampling dates. Gaps in the data are during periods where sensor data were not recorded for one or more rings.", fig.dim=c(7,3), dpi=300}

temp <- left_join(soil, 
                  data.frame(
                    Ring = factor(paste('R', 1:6, sep='')),
                    Trt = factor(c('elevated', 'ambient', 'ambient', 
                                   'elevated', 'elevated', 'ambient'))))
na.dates <- temp %>% 
  filter(is.na(VWC)) %>% 
  select(Date)
temp <- temp %>% 
  mutate(keep = case_when(Date %in% na.dates$Date ~ FALSE, 
                          TRUE ~ TRUE))

## test CO2 effect on soil moisture and temperature
temp[!temp$keep, c('VWC', 'TDRTemp')] <- NA
es <- temp %>% 
  mutate(VWC = asin(sqrt(VWC))) %>% 
  group_by(experiment, Date, Trt) %>% 
  summarise(across(c(VWC, TDRTemp), list(mean=function(x)mean(x, na.rm=TRUE), 
                                         sd=function(x)sd(x, na.rm=TRUE), 
                                         n=function(x)length(x[!is.na(x)])))) %>% 
  mutate(VWC_se = VWC_sd/sqrt(VWC_n), 
         TDRTemp_se = TDRTemp_sd/sqrt(TDRTemp_n)) %>% 
  group_by(experiment, Date) %>% 
  summarise(lrr_vwc = log(VWC_mean[Trt == 'elevated']/VWC_mean[Trt == 'ambient']), 
            se_lrr_vwc = sqrt((VWC_se[Trt == 'elevated']^2/VWC_mean[Trt == 'elevated']^2)+
              (VWC_se[Trt == 'ambient']^2/VWC_mean[Trt == 'ambient']^2)), 
            lrr_temp = log(TDRTemp_mean[Trt == 'elevated']/TDRTemp_mean[Trt == 'ambient']), 
            se_lrr_temp = sqrt((TDRTemp_se[Trt == 'elevated']^2/TDRTemp_mean[Trt == 'elevated']^2)+
              (TDRTemp_se[Trt == 'ambient']^2/TDRTemp_mean[Trt == 'ambient']^2))) %>% 
  mutate(experiment=case_when(experiment == 'expt1' ~ 'First study', 
                              experiment == 'expt2' ~ 'Second study'))

# plot vwc response
ggplot(es, aes(x=Date, y=lrr_vwc)) + 
  geom_line() + 
  geom_ribbon(aes(ymin=lrr_vwc-2*se_lrr_vwc, ymax=lrr_vwc+2*se_lrr_vwc), 
              alpha=0.5, show.legend=FALSE) + 
  labs(x='', y='Log response ratio +/- 95% CI\nlog(VWC [eCO2] / VWC [aCO2])') + 
  facet_grid(cols=vars(experiment), scales='free_x') + 
  geom_vline(xintercept=harvs$Date, linetype='dotted') + 
  geom_hline(yintercept=0)

```

\newpage

```{r, fig.cap='Figure S3. Ordination of litter fungal communities across all characterised samples in the first study. OTU data were analysed using principal coordinates analysis (PCoA) of Bray-Curtis dissimilarities.', fig.dim=c(5,3), dpi=300}

## visualise time and mesh effects

# prepare data
temp <- otus_e1 %>% 
  mutate(harvest=case_when(is.na(Months) ~ 0, 
                           Months==3.2 ~ 3, 
                           Months==6.6 ~ 6, 
                           Months==13 ~ 12))
mat <- decostand(as.matrix(temp %>% select(starts_with('ITSall_'))), 'hellinger')

# ordination
pco <- wcmdscale(vegdist(mat, 'bray', binary=FALSE), eig=TRUE)
pco.scores <- data.frame(temp %>% select(-starts_with('ITSall_')), 
                         scores(pco, display='sites'))
pco.eig <- pco$eig; pco.eig <- pco.eig[pco.eig > 0]; pco.eig <- pco.eig / sum(pco.eig)

# plot result
ggplot(pco.scores, aes(x=Dim1, y=Dim2, colour=as.factor(harvest), shape=Mesh)) + 
  geom_point() + 
  scale_colour_manual(name='Months of decay',
                      values=c(`0`='black', `3`='orange', `6`='red', `12`='purple'),
                      labels=c(`0`='Fresh litter', `3`='Three months',
                               `6`='Six months', `12`='Twelve months')) +
  scale_shape_manual(values=c(initial=1, `2mm`=16, `4mm`=17), 
                     labels=c(initial='Fresh litter', `2mm`='Mesh = 2mm', `4mm`='Mesh = 4mm'), 
                     name='Litter enclosure') + 
  labs(x=paste('PCo 1 (', 100*round(pco.eig[1], 2), ' %)', sep=''), 
       y=paste('PCo 2 (', 100*round(pco.eig[2], 2), ' %)', sep='')) + 
  theme_bw()

```

\newpage

```{r, fig.cap='Figure S4. Ordination of litter fungal communities across all characterised samples in the second study. OTU data were analysed using principal coordinates analysis (PCoA) of Bray-Curtis dissimilarities.', fig.dim=c(5,3), dpi=300}

## visualise time and mesh effects

# prepare data
temp <- otus_e2 %>% 
  mutate(harvest=case_when(is.na(Months) ~ 0, 
                           Months==8 ~ 8, 
                           Months==11.9 ~ 12, 
                           Months==15.6 ~ 16))
mat <- decostand(as.matrix(temp %>% select(starts_with('ITSall_'))), 'hellinger')

# ordination
pco <- wcmdscale(vegdist(mat, 'bray', binary=FALSE), eig=TRUE)
pco.scores <- data.frame(temp %>% select(-starts_with('ITSall_')), 
                         scores(pco, display='sites'))
pco.eig <- pco$eig; pco.eig <- pco.eig[pco.eig > 0]; pco.eig <- pco.eig / sum(pco.eig)

# plot result
ggplot(pco.scores, aes(x=Dim1, y=Dim2, colour=as.factor(harvest))) + 
  geom_point() + 
  scale_colour_manual(name='Months of decay',
                      values=c(`0`='black', `8`='orange', `12`='red', `16`='purple'),
                      labels=c(`0`='Fresh litter', `8`='Eight months',
                               `12`='Twelve months', `16`='Sixteen months')) + 
  labs(x=paste('PCo 1 (', 100*round(pco.eig[1], 2), ' %)', sep=''), 
       y=paste('PCo 2 (', 100*round(pco.eig[2], 2), ' %)', sep='')) + 
  theme_bw()


```

